### 2023-03-02 intro.tex
intro words finished maybe

Local type inference pairs well with type error localization and recovery because information flows systematically down the tree, leading to predictable decisions (as observed also by \cite{pierce}). Languages with constraint-based type inference gather constraints from all relevant locations in a program. This approach is powerful, but it is also notorious for complicating type error localization and recovery, because inconsistencies can arise through a confluence of constraints originating from any number of locations in the program. 
Complex techniques, e.g. based on machine learning or weighting, are needed to heuristically localize errors to expressions. 
In Sec.~\ref{sec:thi}\todo{sec}, we describe how we blend local and constraint-based type inference via type holes in Hazel. In particular, 
we generate constraints for each type hole that appears in the program explicitly or that is  generated internally during the bidirectional type error localization and recovery process. 
We develop a unification algorithm that can recover from inconsistent constraints by maintaining possible type hole fillings. For holes that can be uniquely solved, we suggest the solution. 
If conflicting constraints arise, we localize the error directly to the relevant type holes, rather 
than attempting to localize it to expressions. 
The user can select from different partial fillings, which fills the type hole. From there, the bidirectional system shifts the error to relevant expressions.
Until a choice is made, the system continues to recover gradually, as covered in Sec.~\ref{sec:calculus}.

>>>>

Local type inference pairs well with type error localization and recovery because information flows systematically down the tree, leading to predictable and systematic localization decisions (as informally observed also by \cite{pierce}). Languages with constraint-based type inference gather constraints from all relevant locations in a program. This approach is powerful, but it is also notorious for complicating type error localization and recovery, because inconsistencies can arise through a confluence of constraints originating from any number of locations in the program. 
Complex techniques, e.g. based on machine learning or manual weighting, have been proposed to heuristically localize errors to expressions. 
In Sec.~\ref{sec:thi}\todo{sec}, we describe how we blend local and constraint-based type inference via type holes in Hazel. In particular, 
we generate constraints for each type hole that appears in the program explicitly or that is  generated internally during the bidirectional type error localization and recovery process. 
We develop a unification algorithm that can recover from inconsistent constraints by maintaining possible type hole fillings. For holes that can be uniquely solved, we suggest the solution. 
If conflicting constraints arise, we localize the error directly to the relevant type holes, rather 
than attempting to localize it to expressions. 
The user can select from different partial fillings, which fills the type hole. From there, the bidirectional system shifts the error to relevant expressions.
Until a choice is made, the system continues to recover gradually, as covered in Sec.~\ref{sec:calculus}.
